---
title: "jasa-mandarin_retroflex"
author: "Kate Shan Luo"
date: "7/29/2020"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r}
library(lme4)
library(lmerTest)
library(lsmeans)
library(ggplot2)
library(pastecs)
library(Hmisc)
library(MASS)
library(dplyr)
library(knitr)
library(broom)
library(groupdata2)
library(caret)
library(tidyverse)
library(sjPlot)
library(sjstats)
library(scales)
library(rattle)
library(heplots)
library(klaR)
library(gridExtra)
library(cowplot)
library(multcompView)
library(ggpubr)
library(gss)
library(lme4)
library(lmerTest)
library(lsmeans)
library(ggplot2)
library(pastecs)
library(Hmisc)
library(MASS)
library(dplyr)
library(knitr)
library(broom)
library(groupdata2)
library(caret)
library(tidyverse)
library(sjPlot)
library(sjstats)
library(scales)
library(rattle)
library(heplots)
library(klaR)
library(gridExtra)
library(cowplot)
library(multcompView)
library(ggpubr)
library(gss)
library(extrafont)
library(tinytex)
```

#Table 3
```{r}
data = read.csv("datar2.csv",header=TRUE)
save(data, file = "data.RData")
load(file="data.RData")

table(data$Rating)

data$Rating = factor(data$Rating,
                      levels = c("bunched","retroflex","humped","Alveolar"),
                      labels = c("Bunched","Retroflex","Humped","Alveolar"))
table(data$Rating)

data$Consonant=factor(data$Consonant,
                           levels = c("c", "s", "z","ch","sh","zh"),
                           labels = c("ʦʰ", "s", "ʦ","tʂʰ","ʂ","tʂ"))

data$Grouping = data$Consonant

levels(data$Grouping) <- list(Alveolar=c("ʦʰ","s","ʦ"),Retroflex=c("tʂʰ","ʂ","tʂ"))


data$Participant = factor(data$Participant,
                        levels = c("P1","P2","P3","P4","P5", "P6","P7","P8","P9","P10","P11","P12","P13","P14","P15","P16","P17","P18","P19"),
                        labels = c("P1","P2","P3","P4","P5", "P6","P7","P8","P9","P10","P11","P12","P13","P14","P15","P16","P17","P18","P19"))

data = subset(data, Participant !="P5")
table(data$Rating)

data$Following_Vowel=factor(data$Following_Vowel,
                      levels = c("i", "a", "u"),
                      labels = c("i", "a", "u"))

data$Gender =factor(data$Gender,
                            levels = c("F", "M"),
                            labels = c("F", "M"))

data$F3F2Distance = data$F3_mean - data$F2_mean
```


#Figure 3 articulatory boxplots for different retroflex shapes 
```{r}

dataretroflex = subset(data,Rating != "Alveolar")
save(dataretroflex, file = "dataretroflex.RData")
load(file="dataretroflex.RData")

cleanup = theme(panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.background = element_blank(),
                axis.line = element_line(color = "black"))

cl_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=Curvature_Location))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB"))+
  xlab("")+ylab("Curvature Location")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 0.6,hide.ns = TRUE)+cleanup


cd_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=Curvature_Degree))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB"))+
  xlab("")+ylab("Curvature Degree")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 0.55,hide.ns = TRUE)+ cleanup

MCI_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=MCI))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB"))+
  xlab("")+ylab("MCI")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 4,hide.ns = TRUE)+  cleanup

real1_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=real_1))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB"))+
  xlab("")+ylab("C1_real")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 40,hide.ns = TRUE)+  cleanup


imag1_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=imag_1))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB"))+
  xlab("")+ylab("C1_imag")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 0,hide.ns = TRUE)+  cleanup

fig3 = ggarrange(cl_boxplot,cd_boxplot,MCI_boxplot,
                    real1_boxplot,imag1_boxplot,
                    common.legend = TRUE)

ggsave(fig3,filename= "Fig3_plot.png",dpi = 500, width = 8, height = 6, units = "in")
```

##Figure 4 acoustic boxplots for different retroflex shapes 

```{r}

COG_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=COG))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("COG")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 5000,hide.ns = TRUE)+  cleanup


dispersion_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=Dispersion))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("Dispersion")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 2500,hide.ns = TRUE) + cleanup


skew_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=Skewness))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("Skewness")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 2,hide.ns = TRUE)+  cleanup

F3F2_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=F3F2Distance))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("F3F2Distance")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 1600,hide.ns = TRUE)+  cleanup

F2_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=F2_mean))+
  geom_boxplot(aes(color= Rating))+
  scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("F2")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 2700,hide.ns = TRUE)+ cleanup


F3_boxplot = ggplot(dataretroflex,aes(x=Consonant,y=F3_mean))+
       geom_boxplot(aes(color= Rating))+
    scale_color_manual(values = c("#E7B800","#00AFBB","#00b0f6"))+
  xlab("")+ylab("F3")+
  stat_compare_means(aes(group = Rating),label = "p.signif",label.y = 4000,hide.ns = TRUE)+  cleanup


fig4 = ggarrange(COG_boxplot,dispersion_boxplot,skew_boxplot,
                    F2_boxplot,F3_boxplot,F3F2_boxplot,
                    common.legend = TRUE)

ggsave(fig4,filename= "Fig4_plot.png",dpi = 500, width = 8, height = 6, units = "in")

```


#Figure 5 SSANOVA for all 18 speakers 
```{r}

datassanova = read.csv("datassanova.csv",header=TRUE)

save(datassanova, file = "datassanova.RData")

load(file="datassanova.RData")

datassanova$Section <- NULL 

datassanova$Consonant=factor(datassanova$Consonant,
                      levels = c("c","ch", "s", "sh", "z","zh"),
                      labels = c("ʦʰ","tʂʰ","s", "ʂ", "ʦ","tʂ"))

datassanova$Vowel =factor(datassanova$Vowel,
                          levels = c("i","a","u"),
                          labels = c("i","a","u"))

datassanova$Grouping = datassanova$Consonant

levels(datassanova$Grouping) <- list(Alveolar=c("ʦʰ","s","ʦ"),Retroflex=c("tʂʰ","ʂ","tʂ"))


datassanova$Participant = factor(datassanova$Participant,
                               levels = c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",   "P11","P12","P13","P14","P15","P16","P17","P18","P19"),
                               labels = c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",                             "P11","P12","P13","P14","P15","P16","P17","P18","P19"))

datassanova = subset(datassanova, Participant != "P5")

###p1 tongue ssanove 

p1data = subset(datassanova,Participant == "P1")

p1.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p1data)
p1.tongue.grid <- expand.grid(X = seq(min(p1data$X),max(p1data$X), length = 42), Consonant=levels(p1data$Consonant)) 
p1.tongue.grid$Fit <- predict(p1.tongue, newdata = p1.tongue.grid, se = T)$fit 
p1.tongue.grid$SE <- predict(p1.tongue, newdata = p1.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p1 <- ggplot(p1.tongue.grid,aes(x=X, color=Consonant), res=300)
p1ssanova = comparison.p1 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P1(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position = "none")


##p2 tongue ssanova

p2.tongue = ssanova(Y~Consonant + X + Consonant:X, data=subset(datassanova,Participant =="P2"))
p2.tongue.grid <- expand.grid(X = seq(min(subset(datassanova,Participant =="P2")$X),max(subset(datassanova,Participant =="P2")$X), length = 42), Consonant=levels(data$Consonant)) 
p2.tongue.grid$Fit <- predict(p2.tongue, newdata = p2.tongue.grid, se = T)$fit 
p2.tongue.grid$SE <- predict(p2.tongue, newdata = p2.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p2 <- ggplot(p2.tongue.grid,aes(x=X, color=Consonant), res=300)
p2ssanova = comparison.p2 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P2(M)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")

###p3 ssanova

p3data = subset(datassanova,Participant == "P3")

p3.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p3data)
p3.tongue.grid <- expand.grid(X = seq(min(p3data$X),max(p3data$X),
length = 42), Consonant=levels(p3data$Consonant)) 
p3.tongue.grid$Fit <- predict(p3.tongue, newdata = p3.tongue.grid, se = T)$fit 
p3.tongue.grid$SE <- predict(p3.tongue, newdata = p3.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p3 <- ggplot(p3.tongue.grid,aes(x=X, color=Consonant), res=300)
p3ssanova = comparison.p3 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P3(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


###p4 ssanova

p4data = subset(datassanova,Participant == "P4")

p4.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p4data)
p4.tongue.grid <- expand.grid(X = seq(min(p4data$X),max(p4data$X),
length = 42), Consonant=levels(p4data$Consonant)) 

p4.tongue.grid$Fit <- predict(p4.tongue, newdata = p4.tongue.grid, se = T)$fit 
p4.tongue.grid$SE <- predict(p4.tongue, newdata = p4.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p4 <- ggplot(p4.tongue.grid,aes(x=X, color=Consonant), res=300)
p4ssanova = comparison.p4 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P4(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")

###p6 ssanova

p6data = subset(datassanova,Participant == "P6")

p6.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p6data)

p6.tongue.grid <- expand.grid(X = seq(min(p6data$X),max(p6data$X),
length = 42), Consonant=levels(p6data$Consonant)) 

p6.tongue.grid$Fit <- predict(p6.tongue, newdata = p6.tongue.grid, se = T)$fit 
p6.tongue.grid$SE <- predict(p6.tongue, newdata = p6.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p6 <- ggplot(p6.tongue.grid,aes(x=X, color=Consonant), res=300)
p6ssanova = comparison.p6 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P6(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


###p7 ssanova

p7data = subset(datassanova,Participant == "P7")

p7data <- na.omit(p7data)

p7.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p7data)
p7.tongue.grid <- expand.grid(X = seq(min(p7data$X),max(p7data$X),
                                      length = 42), Consonant=levels(data$Consonant)) 
p7.tongue.grid$Fit <- predict(p7.tongue, newdata = p7.tongue.grid, se = T)$fit 
p7.tongue.grid$SE <- predict(p7.tongue, newdata = p7.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p7 <- ggplot(p7.tongue.grid,aes(x=X, color=Consonant), res=300)
p7ssanova = comparison.p7 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P7(M)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")

###p8 ssanova
p8.tongue = ssanova(Y~Consonant + X + Consonant:X, data=subset(datassanova,Participant =="P8"))
p8.tongue.grid <- expand.grid(X = seq(min(subset(datassanova,Participant =="P8")$X),max(subset(datassanova,Participant =="P8")$X), length  = 42), Consonant=levels(data$Consonant)) 
p8.tongue.grid$Fit <- predict(p8.tongue, newdata = p8.tongue.grid, se = T)$fit 
p8.tongue.grid$SE <- predict(p8.tongue, newdata = p8.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p8 <- ggplot(p8.tongue.grid,aes(x=X, color=Consonant), res=300)
p8ssanova = comparison.p8 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P8(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##p9 ssanova

p9data = subset(datassanova,Participant == "P9")

p9.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p9data)

p9.tongue.grid <- expand.grid(X = seq(min(p9data$X),max(p9data$X),
                                      length = 42), Consonant=levels(p9data$Consonant)) 

p9.tongue.grid$Fit <- predict(p9.tongue, newdata = p9.tongue.grid, se = T)$fit 
p9.tongue.grid$SE <- predict(p9.tongue, newdata = p9.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p9 <- ggplot(p9.tongue.grid,aes(x=X, color=Consonant), res=300)

p9ssanova = comparison.p9 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P9(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##p10 ssanova

p10data = subset(datassanova,Participant == "P10")
p10.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p10data)

p10.tongue.grid <- expand.grid(X = seq(min(p10data$X),max(p10data$X),
 length = 42), Consonant=levels(p10data$Consonant)) 

p10.tongue.grid$Fit <- predict(p10.tongue, newdata = p10.tongue.grid, se = T)$fit 
p10.tongue.grid$SE <- predict(p10.tongue, newdata = p10.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p10 <- ggplot(p10.tongue.grid,aes(x=X, color=Consonant), res=300)
p10ssanova = comparison.p10 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P10(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##p11 ssanova
p11data = subset(datassanova,Participant == "P11")
p11data = na.omit(p11data)

p11.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p11data)

p11.tongue.grid <- expand.grid(X = seq(min(p11data$X),max(p11data$X),
length = 42), Consonant=levels(p11data$Consonant)) 

p11.tongue.grid$Fit <- predict(p11.tongue, newdata = p11.tongue.grid, se = T)$fit 
p11.tongue.grid$SE <- predict(p11.tongue, newdata = p11.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p11 <- ggplot(p11.tongue.grid,aes(x=X, color=Consonant), res=300)
p11ssanova = comparison.p11 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P11(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")



###p12 ssanova
p12data = subset(datassanova,Participant == "P12")
p12data = na.omit(p12data)

p12.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p12data)

p12.tongue.grid <- expand.grid(X = seq(min(p12data$X),max(p12data$X),
length = 42), Consonant=levels(p12data$Consonant)) 

p12.tongue.grid$Fit <- predict(p12.tongue, newdata = p12.tongue.grid, se = T)$fit 
p12.tongue.grid$SE <- predict(p12.tongue, newdata = p12.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p12 <- ggplot(p12.tongue.grid,aes(x=X, color=Consonant), res=300)

p12ssanova = comparison.p12 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P12(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##p13 ssanova
p13data = subset(datassanova,Participant == "P13")

p13data = na.omit(p13data)

p13.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p13data)

p13.tongue.grid <- expand.grid(X = seq(min(p13data$X),max(p13data$X),
length = 42), Consonant=levels(p13data$Consonant)) 

p13.tongue.grid$Fit <- predict(p13.tongue, newdata = p13.tongue.grid, se = T)$fit 
p13.tongue.grid$SE <- predict(p13.tongue, newdata = p13.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p13 <- ggplot(p13.tongue.grid,aes(x=X, color=Consonant), res=300)

p13ssanova = comparison.p13 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P13(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


###p14 ssanova 
p14data = subset(datassanova,Participant == "P14")
p14data = na.omit(p14data)
p14.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p14data)

p14.tongue.grid <- expand.grid(X = seq(min(p14data$X),max(p14data$X),
length = 42), Consonant=levels(p14data$Consonant)) 

p14.tongue.grid$Fit <- predict(p14.tongue, newdata = p14.tongue.grid, se = T)$fit 
p14.tongue.grid$SE <- predict(p14.tongue, newdata = p14.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p14 <- ggplot(p14.tongue.grid,aes(x=X, color=Consonant), res=300)

p14ssanova = comparison.p14 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P14(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##P15 SSANOVA 

p15data = subset(datassanova,Participant == "P15")
p15data = na.omit(p15data)
p15.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p15data)

p15.tongue.grid <- expand.grid(X = seq(min(p15data$X),max(p15data$X),
 length = 42), Consonant=levels(p15data$Consonant)) 

p15.tongue.grid$Fit <- predict(p15.tongue, newdata = p15.tongue.grid, se = T)$fit 
p15.tongue.grid$SE <- predict(p15.tongue, newdata = p15.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p15 <- ggplot(p15.tongue.grid,aes(x=X, color=Consonant), res=300)

p15ssanova = comparison.p15 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P15(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##P16 ssanova
p16data = subset(datassanova,Participant == "P16")
p16data = na.omit(p16data)

p16.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p16data)

p16.tongue.grid <- expand.grid(X = seq(min(p16data$X),max(p16data$X),
length = 42), Consonant=levels(p16data$Consonant)) 

p16.tongue.grid$Fit <- predict(p16.tongue, newdata = p16.tongue.grid, se = T)$fit 
p16.tongue.grid$SE <- predict(p16.tongue, newdata = p16.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p16 <- ggplot(p16.tongue.grid,aes(x=X, color=Consonant), res=300)
p16ssanova = comparison.p16 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P16(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


###P17 ssanova 

p17data = subset(datassanova,Participant == "P17")
p17data = na.omit(p17data)

p17.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p17data)

p17.tongue.grid <- expand.grid(X = seq(min(p17data$X),max(p17data$X),
 length = 42), Consonant=levels(p17data$Consonant)) 

p17.tongue.grid$Fit <- predict(p17.tongue, newdata = p17.tongue.grid, se = T)$fit 
p17.tongue.grid$SE <- predict(p17.tongue, newdata = p17.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p17 <- ggplot(p17.tongue.grid,aes(x=X, color=Consonant), res=300)

p17ssanova = comparison.p17 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P17(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


##P18 ssanova 

p18data = subset(datassanova,Participant == "P18")

p18data = na.omit(p18data)

p18.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p18data)

p18.tongue.grid <- expand.grid(X = seq(min(p18data$X),max(p18data$X),
length = 42), Consonant=levels(p18data$Consonant)) 

p18.tongue.grid$Fit <- predict(p18.tongue, newdata = p18.tongue.grid, se = T)$fit 
p18.tongue.grid$SE <- predict(p18.tongue, newdata = p18.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p18 <- ggplot(p18.tongue.grid,aes(x=X, color=Consonant), res=300)

p18ssanova = comparison.p18 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P18(F)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")

##P19 SSANOVA

p19data = subset(datassanova,Participant == "P19")

p19data = na.omit(p19data)

p19.tongue = ssanova(Y~Consonant + X + Consonant:X, data=p19data)

p19.tongue.grid <- expand.grid(X = seq(min(p19data$X),max(p19data$X),
length = 42), Consonant=levels(p19data$Consonant)) 

p19.tongue.grid$Fit <- predict(p19.tongue, newdata = p19.tongue.grid, se = T)$fit 
p19.tongue.grid$SE <- predict(p19.tongue, newdata = p19.tongue.grid, se = T)$se.fit 

par(pty="s")
comparison.p19 <- ggplot(p19.tongue.grid,aes(x=X, color=Consonant), res=300)

p19ssanova = comparison.p19 + geom_line(aes(y=Fit), size=2, alpha = .8) +
  scale_color_brewer(type="seq", palette = "Paired")+
  ylab("") +xlab("P19(M)")+
  geom_line(aes(y=Fit+(1.96*SE)),lty=2,alpha=0.8)+
  geom_line(aes(y=Fit-(1.96*SE)),lty=2,alpha=0.8)+
  theme(legend.position="none")


fig5 = ggarrange(p1ssanova,p2ssanova,p3ssanova, p4ssanova,
                     p6ssanova,p7ssanova,p8ssanova,p9ssanova,
                     p10ssanova,p11ssanova,p12ssanova,p13ssanova,
                     p14ssanova,p15ssanova,p16ssanova,p17ssanova,
                     p18ssanova,p19ssanova,
                    common.legend = TRUE)

ggsave(fig5,filename= "Fig5_plot.png",dpi = 600, width = 12, height = 12, units = "in")

```



#Tables 3 and 4 LMER 
```{r}

datapolar = read.csv("datapolar.csv",header=TRUE)
save(datapolar, file = "datapolar.RData")
load(file="datapolar.RData")


datapolar$Consonant=factor(datapolar$Consonant,
                     levels = c("c", "s", "z","ch","sh","zh"),
                     labels = c("ʦʰ", "s", "ʦ","tʂʰ","ʂ","tʂ"))

datapolar$Grouping = datapolar$Consonant

levels(datapolar$Grouping) <- list(Alveolar=c("ʦʰ","s","ʦ"),Retroflex=c("tʂʰ","ʂ","tʂ"))

datapolar$Vowel=factor(datapolar$Vowel,
                            levels = c("i", "a", "u"),
                            labels = c("i", "a", "u"))

datapolar$Participant = factor(datapolar$Participant,
                              levels = c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",                            "P11","P12","P13","P14","P15","P16","P17","P18","P19"),
                              labels = c("P1","P2","P3","P4","P5","P6","P7","P8","P9","P10",                            "P11","P12","P13","P14","P15","P16","P17","P18","P19"))

datapolar$Section = factor(datapolar$Section,
                           levels = c("Front","Back"),
                           labels = c("Front","Back"))

datapolar = subset(datapolar, Participant != "P5")
datapolar = na.omit(datapolar)


radiusfront1model = lmer(Radius ~ Grouping * Vowel + (1+Grouping | Participant) + (1+Vowel|Participant), data = subset(datapolar,Section =="Front"),REML= F)

radiusfrontmodel = lmer(Radius ~ Grouping + Vowel + (1+Grouping | Participant) + (1+Vowel|Participant),data = subset(datapolar,Section =="Front"),REML= F)

##Table 3
anova(radiusfront1model,radiusfrontmodel)
summary(radiusfrontmodel)

##Table 4
radiusback1model = lmer(Radius ~ Grouping * Vowel + (1+Grouping | Participant) + (1+Vowel|Participant), data = subset(datapolar, Section =="Back"),REML= F)

radiusbackmodel = lmer(Radius ~ Grouping + Vowel + (1+Grouping | Participant) + (1+Vowel|Participant),data = subset(datapolar, Section =="Back"),REML= F)

anova(radiusback1model, radiusbackmodel)
summary(radiusbackmodel)
```
      


#Figure 6 articulatory values for place-contrast notched boxpltos 
```{r}

cdnotched = ggplot(data,aes(x=Consonant,y=Curvature_Degree,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("Curvature Degree")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 0.5,hide.ns = TRUE)+ cleanup

clnotched = ggplot(data,aes(x=Consonant,y=Curvature_Location,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("Curvature Location")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 0.65,hide.ns = TRUE)+ cleanup


MCInotched= ggplot(data,aes(x=Consonant,y=MCI,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("MCI")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 4,hide.ns = TRUE)+ cleanup

c1realnotched= ggplot(data,aes(x=Consonant,y=real_1,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("C1_real")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 38,hide.ns = TRUE)+  cleanup

C1imagnotched = ggplot(data,aes(x=Consonant,y=imag_1,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("C1_imag")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = -3,hide.ns = TRUE)+  cleanup

fig6 =ggarrange(clnotched,cdnotched,MCInotched,
            c1realnotched,C1imagnotched, 
               common.legend=TRUE)

ggsave(fig6,filename= "Fig6_plot.png",dpi = 500, width = 8, height = 6, units = "in")

```


#Figures 7 and 9, LDA results using articulatory quantifications and acoustic values
```{r}

couleurs=c("darkorange","grey")

Fig7plot = partimat(Grouping ~  real_1 + Curvature_Location, data=data,
                       method="lda",na.action=na.omit, prec=100,display.points=FALSE,image.colors = couleurs,main = "LDA results using Curvature Location and C1_real for all participants", size=12)

dev.copy(png, filename="Figure7_plot.png",res = 500, width = 8, height = 6, units = "in")
dev.off()


fig9plot = partimat(Grouping ~  COG + Dispersion + Skewness + F2_mean,
data=data,method="lda",na.action=na.omit, prec=100,
display.points=FALSE,image.colors = couleurs,
main =  "LDA results using four acoustic values across participants",size = 12)

dev.copy(png,filename="Fig9_plot.png",res = 500, width = 10, height = 8, units = "in")

dev.off()
```


#Figure 8 acoustic values for place contrast notched boxplots  
```{r}
F2notched = ggplot(data,aes(x=Consonant,y=F2_mean,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("F2")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 2800,hide.ns = TRUE)+
  theme(legend.position = "none") +  cleanup


F3notched = ggplot(data,aes(x=Consonant,y=F3_mean,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("F3")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 3800,hide.ns = TRUE)+
  theme(legend.position = "none") + cleanup

F3F2notched = ggplot(data,aes(x=Consonant,y=F3F2Distance,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("F3F2Distance")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 1500,hide.ns = TRUE)+
  theme(legend.position = "none") +  cleanup

COGnotched = ggplot(data,aes(x=Consonant,y=COG,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("COG")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 7500,hide.ns = TRUE)+ cleanup


skewnotched = ggplot(data,aes(x=Consonant,y=Skewness,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("Skewness")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 1.3,hide.ns = TRUE)+
  theme(legend.position = "none") + cleanup


dispersionnotched = ggplot(data,aes(x=Consonant,y=Dispersion,fill=Grouping))+
  geom_boxplot(notch =TRUE)+
  xlab("")+ylab("Dispersion")+
  stat_compare_means(ref.group = "ʦʰ",method = "t.test",label = "p.signif",label.y = 3500,hide.ns = TRUE)+
  theme(legend.position = "none") + cleanup



fig8 = ggarrange( COGnotched, dispersionnotched,skewnotched,
                    F2notched, F3notched, F3F2notched,
                      common.legend = TRUE)

ggsave(fig8,filename= "Fig8_plot.png",dpi = 500, width = 8, height = 6, units = "in")

```



#LDA Results using acoustics 
```{r}
acoustics5 = lda(Grouping ~ COG + Dispersion + Skewness + F2_mean +  F3F2Distance, data=data)

p.acoustics5  = predict(acoustics5,data)
p.acoustics5.class = p.acoustics5$class
p.acoustics5.table = table(p.acoustics5.class,data$Grouping)

acoustics5_accur = sum(diag(p.acoustics5.table))/sum(p.acoustics5.table)

print(acoustics5_accur)


acoustics4 = lda(Grouping ~ COG + Dispersion + Skewness + F2_mean , data=data)

p.acoustics4  = predict(acoustics4,data)
p.acoustics4.class = p.acoustics4$class
p.acoustics4.table = table(p.acoustics4.class,data$Grouping)

acoustics4_accur = sum(diag(p.acoustics4.table))/sum(p.acoustics4.table)

print(acoustics4_accur)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
